---
title: "PSRC Office Search"
output: 
  ioslides_presentation: 
    widescreen: yes
    css: styles.css
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(sf)
library(leaflet)
library(reticulate)
library(data.table)

psrc.logo <- here('data','psrc-logo.png')
trends.logo <- here('data','puget-sound-trends.png')
wgs84 <- 4326
spn <- 2285 

wrk.dir <- "C:/Users/chelmann/OneDrive - Puget Sound Regional Council/coding/gtfs/combined/"
service.period <- c("spring")

# Route IDs by Type
srt_ids <- c('kcm_100340', 'kcm_102638')
lrt_ids <- c('kcm_100479', 'st_100479', 'st_TLINK')
crt_ids <- c('st_SNDR_EV','st_SNDR_TL','st_AMTK', 'st_SNDR_S', 'st_SNDR_N')
brt_ids <- c('ct_701','ct_702','kcm_100512', 'kcm_102548', 'kcm_102576', 'kcm_102581', 'kcm_102615', 'kcm_102619')
pof_ids <- c('kcm_100336', 'kcm_100337','kt_Ferry', 'kt_Annapolis', 'kt_Kitsap Fast Ferry')
fry_ids <- c('wsf_128', 'wsf_145', 'wsf_1621', 'wsf_2022', 'wsf_209', 'wsf_37','wsf_47', 'wsf_229',
             'wsf_2116', 'wsf_2220','wsf_514', 'wsf_73', 'wsf_74', 'wsf_812', 'wsf_920', 'wsf_922')

hct_ids <- c(srt_ids, lrt_ids, crt_ids, brt_ids, pof_ids, fry_ids)

```


```{python, include=FALSE}

import pandas as pd
import transit_service_analyst as tsa
exit
working_directory = 'C:/Users/chelmann/OneDrive - Puget Sound Regional Council/coding/gtfs/combined/'
list_of_years = [('2022','20220330')]

list_of_hct = r.hct_ids

current  = 0
for years in list_of_years:
  path = working_directory + years[0] + '/spring/'
  transit_analyst = tsa.load_gtfs(path, years[1])
  
  # Get Stops that serve High Capacity Transit Routes
  temp = transit_analyst.get_tph_by_line()
  temp = temp[temp['route_id'].isin(list_of_hct)]
  
  route_stops = pd.DataFrame(transit_analyst.get_line_stops_gdf().drop(columns='geometry'))
  route_stops = route_stops[route_stops['rep_trip_id'].isin(temp['rep_trip_id'])]
  unique_stops = route_stops[route_stops['stop_id'].isin(transit_analyst.stops['stop_id'])]
  unique_stops['Year'] = years[0]
  
  if current==0:
    hct_by_route = unique_stops
  else:
    hct_by_route = pd.concat([hct_by_route, unique_stops], ignore_index=True, sort=False)
    
  # Get Route Details for further cleanup before final analysis
  temp = transit_analyst.routes[['agency_id','route_id','route_short_name','route_long_name','route_type']]
  temp['Year'] = years[0]
  
  if current==0:
    routes = temp
  else:
    routes = pd.concat([routes, temp], ignore_index=True, sort=False)
    
  current = current + 1
  
# end of loop
exit
```

```{r route-stops-cleanup, include=FALSE}

stops <- as_tibble(fread(paste0(wrk.dir,"2022/spring/stops.txt"))) %>% 
  select(stop_id,stop_lat,stop_lon)

stops.layer <- st_as_sf(stops, coords = c("stop_lon","stop_lat"), crs = wgs84)

hct.route.stops <- as_tibble(py$hct_by_route) %>% 
  select(stop_id, route_id, Year) %>% 
  distinct() %>%
  mutate(Type = case_when(
    route_id %in% srt_ids ~ "Streetcar",
    route_id %in% lrt_ids ~ "Light Rail",
    route_id %in% crt_ids ~ "Commuter Rail",
    route_id %in% brt_ids ~ "Bus Rapid Transit",
    route_id %in% pof_ids ~ "Ferry",
    route_id %in% fry_ids ~ "Ferry"))

hct.stops <- hct.route.stops %>% select(stop_id) %>% pull() %>% unique()
hct.stops.layer <- stops.layer %>% filter(stop_id %in% hct.stops)
hct.stops.layer <- left_join(hct.stops.layer, hct.route.stops, by=c("stop_id"))

```

```{r map-creation, include=FALSE}
psrc.zipcodes <- st_read(here('data', 'psrc_zipcodes.shp')) %>% st_transform(wgs84) %>% select(zipcode)
psrc.zipcodes.centroids <- psrc.zipcodes %>% st_transform(spn) %>% st_centroid() %>% st_transform(wgs84)

zipcode.data <- read_csv(here('data', 'staff-zipcodes.csv')) %>% mutate(response=1, Zip=as.character(Zip)) 
zipcode.point.response <- left_join(psrc.zipcodes.centroids, zipcode.data, by=c("zipcode"="Zip")) %>% drop_na()

building.data <- read_csv(here('data', 'buildings.csv'))
building.data <- st_as_sf(building.data, coords = c("Lon","Lat"))

i <- awesomeIcons(icon = 'user',iconColor = 'black',library = 'fa')

# Building Layer
getColor <- function(building.data) {
  sapply(building.data$Code, function(Code) {
  if(Code == 0) {
    "orange"
  } else if(Code == 1) {
    "blue"
  } else {
    "green"
  } })
}

o <- awesomeIcons(icon = 'building',iconColor = 'black', library = 'fa', markerColor = getColor(building.data))

routeCol <- colorFactor(palette = 'RdYlGn', hct.stops.layer$Type)

m <- leaflet() %>% 
        addProviderTiles(providers$CartoDB.Positron) %>%
        
        addLayersControl(baseGroups = c("Base Map"),
                         overlayGroups = c("Zipcode","Responses","Transit Stops"),
                         options = layersControlOptions(collapsed = TRUE)) %>%
        
        addEasyButton(easyButton(
            icon="fa-globe", title="Region",
            onClick=JS("function(btn, map){map.setView([47.615,-122.257],8.5); }"))) %>%
        
        addPolygons(data=psrc.zipcodes,
                    fillOpacity = 0.0,
                    opacity = 1.0,
                    weight = 1,
                    color = "#91268F",
                    dashArray = "1",
                    group = "Zipcode") %>%
        
        addAwesomeMarkers(data=zipcode.point.response, 
                          icon = i, 
                          clusterOptions = markerClusterOptions(),
                          group="Responses") %>%
  
        addCircles(data=hct.stops.layer, 
                          group="Transit Stops",
                          color = ~routeCol(Type),
                          label= ~as.character(Type)) %>%
  
        addLegend(pal = routeCol,
                  values = hct.stops.layer$Type,
                  group = "Transit Stops",
                  position = "bottomleft",
                  title = "HCT Transit Stops") %>%
  
        hideGroup("Transit Stops")

b <- leaflet() %>% 
        addProviderTiles(providers$CartoDB.Positron) %>%
        
        addLayersControl(baseGroups = c("Base Map"),
                         overlayGroups = c("Buildings","Transit Stops"),
                         options = layersControlOptions(collapsed = TRUE)) %>%
        
        addEasyButton(easyButton(
            icon="fa-building", title="Office Building Locations",
            onClick=JS("function(btn, map){map.setView([47.606327,-122.334673],15); }"))) %>%
  
        setView(lat=47.606327, lng=-122.334673, zoom=15) %>%
  
        addAwesomeMarkers(data=building.data, 
                          icon = o, 
                          group="Buildings",
                          label=~as.character(Building)) %>%
  
        addCircles(data=hct.stops.layer, 
                          group="Transit Stops",
                          color = ~routeCol(Type),
                          label= ~as.character(Type)) %>%
  
        addLegend(pal = routeCol,
                  values = hct.stops.layer$Type,
                  group = "Transit Stops",
                  position = "bottomleft",
                  title = "HCT Transit Stops")

```


## Home{.slide_body_text_sm}
<div style="position:absolute; top:5%; left:5%;">
<slideheadstyle>Home Zipcode Locations</slideheadstyle>
</div>

<div style="float: left; width: 25%;">
* 63 total staff in this summary
* Staff are located in 3 out of 4 PSRC counties
* 3 staff are currently located in Kitsap County and are ferry dependent for in-person events
* 6 staff members live in Snohomish County
</div> 

<div style="float: right; width: 75%;">
`r m`
</div> 

<IMG style="position:absolute; bottom:2.5%; left:5%; width:192px; "SRC="`r psrc.logo`">


## Office{.slide_body_text_sm}
<div style="position:absolute; top:5%; left:5%;">
<slideheadstyle>Office Search Locations</slideheadstyle>
</div>

<div style="float: left; width: 25%;">
* Explored 10 buildings in the Seattle Central Business District
* All locations are within a few blocks of High-Capacity Transit
</div> 

<div style="float: right; width: 75%;">
`r b`
</div> 

<IMG style="position:absolute; bottom:2.5%; left:5%; width:192px; "SRC="`r psrc.logo`">


## Questions {data-background=images/rhs-slide-2400x800.jpg data-background-size=cover}

<div style="position:absolute; top:80%; left:5%;">
<transitionstyle>Questions? <br></transitionstyle>
<contactstyle>NBGrennan@psrc.org or CHelmann@psrc.org</contactstyle>
</div>